---
title: "Assignment(FML)"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2023-09-10"
---
#data collected from Kaggle("https://www.kaggle.com/datasets/benroshan/factors-affecting-campus-placement")
```{r}
c<-read.csv("E:\\campus\\Placement_Data_Full_Class.csv")
c
```

#Summary for all the individual columns
```{r}
summary(c)
```
#Descriptive statistics of Quantitative variables ('mean','mode','median','standard_deviation,'variance','sum','minimum','maximum')
```{r}
mean(c$ssc_p)
```
```{r}
median(c$hsc_p)
```
```{r}
mode(c$degree_p)
```
```{r}
sd(c$ssc_p)
```
```{r}
var(c$hsc_p)
```
```{r}
sum(c$degree_p)
```
```{r}
min(c$etest_p)
max(c$etest_p)
```
#Descriptive statistics for Quanlitative/Categorial variables
```{r}
mode(c$ssc_b)
```
```{r}
table(c$gender)
```
```{r}
table(c$ssc_b)
```
```{r}
table(c$hsc_b)
```
#Transform the quantitative variables
```{r}
log(c$salary)
```
```{r}
f<-(c$etest_p)-mean(c$etest_p)/sd(c$etest_p)
f
```
#Plotting the quantitative variable
```{r}
barplot(c$etest_p,c$degree_p,main="Bar plot")
```
```{r}
hist(c$ssc_p,main="Histogram")
```
#scatter plot for quantitative variables
```{r}
plot(c$ssc_p,c$hsc_p,main="Scatter Plot")
```

